---
title: "Lab 3"
author: "Echo Tang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Read in data

```{r Read in data, results=FALSE}
library(tidyverse)
library(data.table)

download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")

```

## Question 2: Dimensions, headers, footers

```{r Dimensions, headers, footers}
dim(met)
head(met)
tail(met)
```
There are 2377343 observations and 30 variables in this dataset. 

## Question 3: Take a look at variables

```{r Take a look at variables}
str(met)

```

## Question 4: Looking at key variables

```{r Take a look at key variables}
table(met$year)
table(met$day)
table(met$hour)

summary(met$temp)
summary(met$elev)
summary(met$wind.sp)

met[met$elev==9999.0] <- NA
summary(met$elev)

# The highest elevation is 4113 meters.

met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)

```

## Question 5: Checking with external sources

According to the met data table, the minimum temperature of -40 degrees Celsius is found in Louisiana, which does not make sense for a temperature in August. Other than that observation, the temperatures are higher and will be cleaned accordingly. 

```{r Checking against external sources}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)

```
The new dataset's minimum temperature is -3 degrees Celsius, which makes more sense given the month. 

## Question 6: Summary statistics

```{r Summary statistics}
elev <- met[elev==max(elev)]
summary(elev)

cor(elev$temp, elev$wind.sp, use="complete")
cor(elev$temp, elev$hour, use="complete")
cor(elev$wind.sp, elev$day, use="complete")
cor(elev$wind.sp, elev$hour, use="complete")
cor(elev$temp, elev$day, use="complete")

```
## Question 7: Exploratory graphs

```{r Exploratory graphs}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)

```
```{r Exploratory graph continued}
library(leaflet)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)

```

```{r Making time series}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(date)]
head(elev)

```

```{r Plotting time series}
plot(elev$date, elev$temp, type='l')
plot(elev$date, elev$wind.sp, type='l')

```

Some trends are that oftentimes for higher temperatures, the wind speed is lower. For example, on August 7, the wind speed has a low peak; the corresponding temperature is a high peak in the temperature time-series plot. Similarly, between August 22 and 27, the wind speeds are higher than other parts of the graph. Correspondingly, there are some low peaks of temperatures for dates between August 22 and 27. However, compared to the temperature graph where there are small moments of "plateauing," the wind speeds seem to have more variation in their observations. 


